---
title: "R_prgramming"
author: "Yue"
date: "15/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Object

    R has 5 basic atomic class of objects: 

-  character
-  numeric : 1 , Inf(Infinity) , NaN( undefined number, e.g. 0/0)
-  integer : 1L
-  complex
-  logical(True or False)
 
 
    The most basic object is a vector:
    
-  A vector only contain objects of the same class
-  Exception : *list* which represented as a vector but can contain different classes.

    Empty vector can be created with vector() fun function,
    
## Attibutes 

-  names, dimmanes
-  dimensions ( num of rows and num of columns)
-  class
-  length
-  other user-defined attributes

can be accessed using attributes() function


## Creating Vectors

using the c() function
```{}
x<-c(0.5, 0.6)    ## numeric
x<-c(TRUE,FALSE)  ## logical
x<-c(T,F)         ## logical
x<-c("a","b","c") ## character
x<- 9;29          ## integer
x<-c(1+0i,2+4i)   ## complex

```


Using the vector() function
```{}
x<-vector("numeric", length=10)
x
```

When different objects are mixed in a vector, *coercion* occurs so that every element in the vector is of the same class( usually the least common one)
```{r}
y<-c(1.7,"a")    ## character
y<-c(TRUE,2)     ## numeric

```


